---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r 1}
unzip("activity.zip")
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date, "%Y-%m-%d")
library(dplyr)
```

## What is mean total number of steps taken per day?
```{r 2&3}
stepsbyday <- activity %>% 
      group_by(date) %>% 
      summarise(totalSteps = sum(steps, na.rm=TRUE))

with(stepsbyday, plot(date, totalSteps, 
                      type = "h",
                      main = "Total number of steps taken each day",
                      xlab = "Date",
                      ylab = "Total number of steps"))
```
  
#### Mean number of steps taken each day = `r mean(stepsbyday$totalSteps)`  
#### Median number of steps taken each day = `r median(stepsbyday$totalSteps)`

## What is the average daily activity pattern?
```{r 4&5}
stepsbytime <- activity %>% 
      group_by(interval) %>% 
      summarise(avgSteps = mean(steps, na.rm=TRUE))

with(stepsbytime, plot(interval, avgSteps, 
                       type="l",
                       main = "Time series plot of the average number of steps taken",
                       xlab = "Interval",
                       ylab = "Average number of steps"))

maxStepTime <- stepsbytime[stepsbytime$avgSteps == max(stepsbytime$avgSteps),1]
```
  
#### The 5-minute interval that, on average, contains the maximum number of steps = `r maxStepTime`

## Imputing missing values

#### Total number of **NA** in the dataset = `r sum(activity$steps %in% NA)`
```{r 6&7}
## Create a new dataset and replace NA with mean for that 5-minute interval
imputed <- merge(activity, stepsbytime)
imputed[imputed$steps %in% NA, 2] <- imputed[imputed$steps %in% NA, 4] 

newstepsbyday <- imputed %>% 
          group_by(date) %>% 
          summarise(totalSteps = sum(steps, na.rm=TRUE))

with(newstepsbyday, plot(date, totalSteps, 
                      type = "h",
                      main = "Total number of steps taken each day after missing values are imputed",
                      xlab = "Date",
                      ylab = "Total number of steps"))
```
  
#### Mean number of steps taken each day = `r mean(newstepsbyday$totalSteps)`  
#### Median number of steps taken each day = `r median(newstepsbyday$totalSteps)`


## Are there differences in activity patterns between weekdays and weekends?
``` {r 8}
# create a vector of weekdays
wdays <- c("Monday", "Tuesday", "Wednesay", "Thursday", "Friday")
# Use `%in%` and `weekdays` to create a logical vector
# convert to `factor` and specify the `levels/labels`
imputed$wDay <- factor((weekdays(imputed$date) %in% wdays), 
                       levels = c(FALSE, TRUE), 
                       labels = c("weekend", "weekday"))

newstepsbytime <- imputed %>% 
      group_by(interval, wDay) %>% 
      summarise(avgSteps = mean(steps, na.rm=TRUE))

library(lattice)
xyplot(avgSteps ~ interval | wDay, data = newstepsbytime, 
       main = "Time series plot of the average number of steps taken",
       xlab = "Interval",
       ylab = "Average number of steps",
       layout = c(1,2),
       type="l")

```
